\textls[-0]{\thetitle}

### 2022-11-11 metatheory-condensed.tex
edits in FMLTT sec

In rule \ruleref{l/add}, the second premise $\goodTerm{\Gm, A}{t}{T}$ is
responsible for late binding.
Here, $A$ abstracts the fields preceding the current field~$t$
(i.e., those recorded by the first premise $\goodTerm{\Gm}{\lkg}{\TyLkg{\lsig}}$).
Crucially, %unlike the second field in a dependent pair (\ruleref{tm/pair}),
the premise $\goodTerm{\Gm, A}{t}{T}$ makes clear that
the typing of~$t$ in \ruleref{l/add} is \emph{universally} quantified over
how the fields in $t$'s context are defined.
So a different linkage~$\ell'$ that overrides fields in $t$'s context (and thus models a derived family)
can reuse~$t$---without retyping it---by first
projecting~$t$ from $\LkgAdd{\lkg}{t}$ and then appending it to~$\lkg'$.
%\EDJ{I suggest change this sentence to `by projecting it out of parent linkage and adding it in the derived linkage'}.

This type~$A$ abstracting the context in \ruleref{l/add} and
\ruleref{lsig/add} do not necessarily contain the same types of the prior fields as recorded by~$\TyLkg{\lsig}$,
%\EDJ{Suggest changing to ``not necessarily the same as those recorded by lsig in the prior fields''},
because a field representing the code $\wcode{\wsig}$ of a W-type has to
expose different types to different fields that come after it.
Later fields that invoke
$\Wsup{\wsig}{\cDot}{\cDot}$ or $\Wrec{\wsig}{\cDot}{\cDot}$
should see the concrete signature~$\wsig$, as rules \ruleref{tm/wsup}
and \ruleref{tm/wrec} stipulate.
By contrast, $\wsig$ should be hidden
from all other fields, so that they can be reused
in a different context where the W-type signature~$\wsig$ is replaced by an extended one~$\wsig'$.

@@

We illustrate this with \cref{fig:stlc-linkage-typing}.
On the left is code excerpted from \cref{fig:stlc-mechanized}, and on the right
is how the corresponding fields are modeled and typed in \TT.
%
$\lkg_i$ is the linkage that adds the $i$-th field~$t_i$ with the typing
$\goodTerm{%
\tikzmarkin[disable rounded corners=true,set fill color=A-color,set border color=A-color]{A-i}(-0.02,-0.10)(-0.02,0.28)
A_i%
\tikzmarkend{A-i}%
}{%
\tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t-i}(-0.02,-0.10)(-0.02,0.28)%
t_i%
\tikzmarkend{t-i}%
}{%
\tikzmarkin[disable rounded corners=true,set fill color=T-color,set border color=T-color]{T-i}(-0.02,-0.10)(-0.02,0.28)%
T_i%
\tikzmarkend{T-i}%
}$.
%
For readability, we feel free to use a notation with explicit names
in this typing judgment:
we name the context as $\mathit{self}$ (cf.~\cref{fig:stlc-compiled})
and feel free to write its type~$A_i$, which is a dependent tuple type, as a
dependent record type with labels.
%
\begin{itemize}
% [leftmargin=4ex,itemsep=.5ex]
  [labelsep=*,leftmargin=1pc,itemsep=3pt]

>>>>

In rule \ruleref{l/add}, the second premise $\goodTerm{\Gm, A}{t}{T}$ is
responsible for late binding.
Here, $A$ abstracts the context of the current field~$t$, controlling
how the types of the fields prior to~$t$ are exposed to the typing of~$t$.
Crucially, %unlike the second field in a dependent pair (\ruleref{tm/pair}),
the premise $\goodTerm{\Gm, A}{t}{T}$ makes clear that
the typing of~$t$ in \ruleref{l/add} is \emph{universally} quantified---%
rather than \emph{existentially} quantified as is in \ruleref{tm/pair}---%
over how the fields in $t$'s context are defined.
Late binding enables reuse.
A different linkage~$\ell'$ that overrides fields in $t$'s context (and thus models a derived family)
can reuse~$t$---without retyping it---by first
projecting~$t$ from $\LkgAdd{\lkg}{t}$ and then appending it to~$\lkg'$.
%\EDJ{I suggest change this sentence to `by projecting it out of parent linkage and adding it in the derived linkage'}.

%recorded by the first premise $\goodTerm{\Gm}{\lkg}{\TyLkg{\lsig}}$
The type~$A$ in \ruleref{l/add} and \ruleref{lsig/add},
abstracting the types of the prior fields, does not necessarily contain the same types as those recorded by~$\TyLkg{\lsig}$,
%\EDJ{Suggest changing to ``not necessarily the same as those recorded by lsig in the prior fields''},
because a field defining the code $\wcode{\wsig}$ of a W-type has to
expose different types to different fields that come after it.
Later fields that invoke
$\Wsup{\wsig}{\cDot}{\cDot}$ or $\Wrec{\wsig}{\cDot}{\cDot}$
should see the concrete signature~$\wsig$, as rules \ruleref{tm/wsup}
and \ruleref{tm/wrec} stipulate.
By contrast, $\wsig$ should be hidden
from all other fields, so that they can be reused
in a different context where the W-type signature~$\wsig$ is replaced by an extended one~$\wsig'$.

@@

We use \cref{fig:stlc-linkage-typing} to illustrate.
On the left is code excerpted from \cref{fig:stlc-mechanized}, and on the right
is how the corresponding fields are modeled and typed in \TT.
%
$\lkg_i$ is the linkage that adds the $i$-th field~$t_i$ with the typing
$\goodTerm{%
\tikzmarkin[disable rounded corners=true,set fill color=A-color,set border color=A-color]{A-i}(-0.02,-0.10)(-0.02,0.28)
A_i%
\tikzmarkend{A-i}%
}{%
\tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t-i}(-0.02,-0.10)(-0.02,0.28)%
t_i%
\tikzmarkend{t-i}%
}{%
\tikzmarkin[disable rounded corners=true,set fill color=T-color,set border color=T-color]{T-i}(-0.02,-0.10)(-0.02,0.28)%
T_i%
\tikzmarkend{T-i}%
}$.
%
For readability, we feel free to use a notation with explicit names
in this typing judgment:
we name the context as $\mathit{self}$ (cf.~\cref{fig:stlc-compiled})
and feel free to write its type~$A_i$, which is a dependent tuple type, as a
dependent record type with labels.
%
\begin{itemize}
% [leftmargin=4ex,itemsep=.5ex]
  [labelsep=*,leftmargin=1pc,itemsep=3pt]

### 2022-11-11 stlc-linkage-typing.tex
columnize
edits in FMLTT sec

\begin{tabular}{@{}l@{\ \ \ }L@{}}

@@

\quad

@@

\quad

@@

\quad
\quad

@@

@@

\caption{\TT encoding of the \texttt{STLC} family from \cref{fig:stlc-mechanized}.}
\label{fig:stlc-linkage-typing}
\end{figure}

>>>>

\begin{tabular}{@{}l@{\ \ \ }L@{\quad}L@{\quad}L@{\quad}L@{}}

@@

@@

\goodTerm{\cdot}{\lkg_0}{\TyLkg{\lsig_0}}

@@

@@

\multicolumn{4}{@{}l@{}}{

@@

\caption{\TT encoding of the \texttt{STLC} family from \cref{fig:stlc-mechanized}.
Each row $\goodTerm{\mathit{self}\!:\!
\tikzmarkin[disable rounded corners=true,set fill color=A-color,set border color=A-color]{A-i-cap}(-0.02,-0.10)(-0.02,0.28)
A_i%
\tikzmarkend{A-i-cap}%
}{%
\tikzmarkin[disable rounded corners=true,set fill color=t-color,set border color=t-color]{t-i-cap}(-0.02,-0.10)(-0.02,0.28)%
t_i%
\tikzmarkend{t-i-cap}%
}{%
\tikzmarkin[disable rounded corners=true,set fill color=T-color,set border color=T-color]{T-i-cap}(-0.02,-0.10)(-0.02,0.28)%
T_i%
\tikzmarkend{T-i-cap}%
}$ types a field.
The type $A_i$ controls how the typing of field $t_i$ sees the types of the fields prior to $t_i$.
}
\label{fig:stlc-linkage-typing}
\end{figure}
